---
title: "Human_NucSeq_Plots"
author: "Georgina Dowsett"
date: '`r Sys.Date()`'
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, root.dir = '~/Dropbox/Thesis/Chapter3/')
```


# Plotting of snRNA-seq data from the human hypothalamus for Thesis Figures

This R Markdown document provides the code required for plots of the snRNA-seq datset of human hypothalamus

```{r message = FALSE}
library(Seurat)
library(ggplot2)
library(tidyverse) 
library(scCustomize)
library(patchwork)
library(RColorBrewer)
library(gridExtra)
library(viridis)
library(viridisLite)
library(scales)
library(pheatmap)
source('~/Dropbox/Data/ST_forhumanhypomap/Scripts/RNAseqSTUtils.R')

```

```{r echo = FALSE}
sessionInfo()
```

## Load the data and set output folder 
```{r, eval = FALSE}
data<-readRDS('~/Dropbox/Data/ST_forhumanhypomap/data/NucSeq/human_hypo_combined.rds')
markers<-read.delim('~/Dropbox/Data/ST_forhumanhypomap/data/markers_final.txt')
num_to_name<-read.delim('~/Dropbox/Data/ST_forhumanhypomap/data/NucSeq/human_hypo_combined_annotation_overview.txt')
Idents(data)<-data@meta.data$C6
dim(data)
results_folder<-'~/Dropbox/Thesis/Chapter3/Figs/'

```



## Overview of the NucSeq dataset. 

First, we need to annotate the clusters according to general cell types, and add cluster names to the object 
```{r, eval = FALSE} 
c4<-data.frame(table(data$C4))
c4 <- c4 %>%
  mutate(Var1 = as.character(Var1), cell_type = ifelse(Var1 %in% c("C4-116", "C4-117", "C4-118", "C4-119", "C4-120", "C4-121", "C4-122", "C4-123", "C4-124", "C4-125", "C4-126", "C4-127", "C4-128"), "Astrocytes",
                                                       ifelse(Var1 %in% c("C3-54"), "OPCs",
                                                              ifelse(Var1 %in% c("C3-55", "C3-56", "C3-57", "C3-58", "C3-59"), "Oligos",
                                                                     ifelse(grepl("^C4-(1|2|3|4|5|6|7|8|9|10|11|12|13|14|15|16|17|18|19|20|21|22|23|24|25|26|27|28|29|30|31|32|33|34|35|36|37|38|39|40|41|42|43|44|45|46|47|48|49|50|51|52|53|54|55|56|57|58|59|60|61|62|63|64|65|66|67|68|69|70|71|72|73|74|75|76|77|78|79|80|81|82|83|84|85|86|87|88|89|90|91|92|93)$", Var1), "Neurons",
                                                                            ifelse(Var1 == "C4-94", "Ependymal",
                                                                                   ifelse(Var1 == "C4-95", "Tanycytes",
                                                                                          ifelse(Var1 %in% c("C4-96", "C4-97"), "Ependymal",
                                                                                                 ifelse(Var1 %in% c("C4-98", "C4-99"), "Endothelial",
                                                                                                        ifelse(Var1 %in% c("C4-100", "C4-101", "C4-102"), "Mural",
                                                                                                               ifelse(Var1 %in% c("C4-103", "C4-104"), "Ependymal",
                                                                                                                      ifelse(Var1 %in% c("C4-105", "C4-106", "C4-107", "C4-108", "C4-109", "C4-110", "C4-111", "C4-112", "C4-113", "C4-114", "C4-115"), "Immune",
                                                                                                                             "unknown"
                                                                                                                      )
                                                                                                               )
                                                                                                        )
                                                                                                 )
                                                                                          )
                                                                                   )
                                                                            )
                                                                     )
                                                              )
                                                       )
  )
  )
c4$Var1 <- as.character(c4$Var1)
c4$cell_type <- as.character(c4$cell_type)
metadata_map <- setNames(c4$cell_type, c4$Var1)
c4_var1 <- unlist(data[["C4"]])
data[["cell_type"]] <- metadata_map[c4_var1]

cluster_levels <- c("C1", "C2", "C3", "C4", "C5", "C6")
for (level in cluster_levels) {
  sub_num_to_name <- num_to_name[num_to_name$cluster %in% unique(data[[level]])[,level], ]
  cluster_mapping <- setNames(sub_num_to_name$genes_conc, sub_num_to_name$cluster)
  cluster_names <- cluster_mapping[match(data[[level]][,level], names(cluster_mapping))]
  data <- AddMetaData(data, metadata = cluster_names, col.name = paste0(level, "_names"))
}
```

## Fig 3-1
```{r, eval = FALSE}
DimPlot_scCustom(data, raster = F, group.by = 'cell_type', reduction = 'umap_scvi_hypo')
print(FeaturePlot(data, c('SNAP25', 'AQP4', 'PLP1', 'PDGFRA', 'FYB1', 'VIM'), order = F, reduction = 'umap_scvi_hypo', raster = F, ncol = 3, cols = c('lightgrey', 'darkred'), pt.size =0.1) &
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank()))
```

## Fig 3-2
We look at expression of transcription factors, expressed in different regions of the hypothalamus. 

```{r, eval = FALSE}
print(FeaturePlot(data, c('MEIS2', 'LHX6', 'FEZF1', 'TBX3', 'SIM1', 'FOXB1', 'LMX1A', 'PITX2'), order = T, reduction = 'umap_scvi_hypo', raster = F, ncol = 2, cols = c('lightgrey', 'darkblue'), pt.size =0.1) &
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank()))
```

## Fig 3-3
We show umap plots coloured by dataset and by donor 

```{r, eval = FALSE}
DimPlot_scCustom(data, group.by = 'Dataset', reduction = 'umap_scvi_hypo', raster = F, colors_use = c('goldenrod3', 'darkgreen')) & 
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank())
DimPlot_scCustom(data, group.by = 'Donor_ID', reduction = 'umap_scvi_hypo', raster = F) & 
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank())

```

NB Fig 3-4 is taken from the human HYPOMAP paper

## Fig 3-5
```{r, eval = FALSE}
Idents(data)<-data@meta.data$C3
arc<-WhichCells(data, idents = c('C3-22'))
data$ARC <- ifelse(colnames(data) %in% arc, 'ARC', 'NO')
cols = c('ARC' = 'purple', 'NO' = 'lightgrey')
DimPlot(data, group.by = 'ARC', cols = cols, reduction = 'umap_scvi_hypo', raster = FALSE)

arc<-subset(data, subset = C3 == 'C3-22')
Idents(arc)<-arc@meta.data$C6
arc_markers<-FindAllMarkers(arc, only.pos = TRUE)
arc_markers$specificity_score <- ((arc_markers$pct.1 + 0.01) / (arc_markers$pct.2 + 0.01)) * arc_markers$avg_log2FC

#First we want to add new names which are the same as C6_names, minus the first 16 characters
names<-data.frame(table(arc$C6_names))
names$new_names <- substring(names$Var1, 17)
new_names<- c( "NE_Mid-1_3_TBX3_ESR1_CALCR" = "ESR1_CALCR",
               "NE_Mid-1_3_TBX3_ESR1_CCBE1_COL22A1" = "ESR1_CCBE1_COL22A1",
               "NE_Mid-1_3_TBX3_ESR1_CCBE1_IGF1" = "ESR1_CCBE1_IGF1",
               "NE_Mid-1_3_TBX3_ESR1_COL22A1" = "ESR1_COL22A1",
               "NE_Mid-1_3_TBX3_ESR1_PGR_CALCR" = "ESR1_PGR_CALCR",
               "NE_Mid-1_3_TBX3_ESR1_PGR_LYPD6B" = "ESR1_PGR_LYPD6B",
               "NE_Mid-1_3_TBX3_PDGFD_GABRE_ALDH1A1" = "PDGFD_GABRE_ALDH1A1", 
               "NE_Mid-1_3_TBX3_PDGFD_GABRE_PGR" = "PDGFD_GABRE_PGR",
               "NE_Mid-1_3_TBX3_PDGFD_GABRE_TWIST1" = "PDGFD_GABRE_TWIST1",
               "NE_Mid-1_3_TBX3_PDGFD_PIK3C2G_HMCN1" = "PDGFD_PIK3C2G_HMCN1",
               "NE_Mid-1_3_TBX3_PDGFD_PIK3C2G_WIF1" = "PDGFD_PIK3C2G_WIF1",
               "NE_Mid-1_3_TBX3_POMC_ANKRD30A" = "POMC_ANKRD30A",
               "NE_Mid-1_3_TBX3_POMC_CALCR" = "POMC_CALCR",
               "NE_Mid-1_3_TBX3_POMC_LEPR" = "POMC_LEPR")
arc$naming <- new_names[as.character(arc@meta.data$C6_names)]

DimPlot(arc, group.by = 'naming', label = T, reduction = 'umap_scvi_hypo', raster = F, repel = T) & 
  NoLegend() &
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank()) &
  xlim(7, 14) &
  ylim(-7, 0)

genes_of_interest <- c('TBX3', 'POMC', 'PDGFD', 'ESR1', 'LEPR', 'CALCR', 'ANKRD30A', 'PIK3C2G', 'WIF1', 'HMCN1', 'GABRE', 'TWIST1', 'ALDH1A1', 'PGR', 'LYPD6B', 'COL22A1', 'CCBE1', 'IGF1', 'SLC17A6', 'SLC32A1')

arc_percents<-calc_percent_cluster(arc, genes_of_interest, 'C6_names')
pheatmap::pheatmap(arc_percents[1:20], color = viridis(256), border_color = 'lightgrey', cluster_rows = F, cluster_cols = F, fontsize = 10)
```

## Fig 2-14
```{r, eval = FALSE}
DimPlot_scCustom(data, reduction = 'umap_scvi_hypo', raster = F, group.by = 'C1_names', label = T) & NoLegend()
OL<-subset(data, subset = C1_names == c('Oligo_Mature', 'Oligo_Precursor'))
pdf(paste0(results_folder, 'Oligo-C1-names.pdf'), width = 3, height = 6)
DimPlot(OL, reduction = 'umap_scvi_hypo', raster = F, group.by = 'C1_names', label = T) & 
  NoLegend() &
  xlim(-10, -4) &
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank())
DimPlot_scCustom(OL, reduction = 'umap_scvi_hypo', raster = F, group.by = 'C2_names', label = T, repel = T) & 
  NoLegend() &
  xlim(-10, -4) &
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank())
DimPlot_scCustom(OL, reduction = 'umap_scvi_hypo', raster = F, group.by = 'C3_names', label = T, repel = T) & 
  NoLegend() &
  xlim(-10, -4) &
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank())
dev.off()
FeaturePlot(OL, reduction = 'umap_scvi_hypo', raster = F, features = c('PDGFRA', 'CSPG4', 'BMP4', 'OLIG2', 'MBP', 'MOG'), ncol = 3) & 
  xlim(-10, -4) &
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank())
```

## Fig 3-16
Calculating the number of cells per cluster and also the % donor & dataset as well as proportion donor & dataset
```{r, eval = FALSE}
print(table(OL$C1_names))
print(table(OL$C2_names))
dataset<-data.frame(table(OL$C2_names, OL$Dataset))
dataset<-pivot_wider(dataset, names_from = Var1, values_from = Freq)
x<-data.frame(table(data$Dataset))
dataset<-merge(dataset, x, by.y = 'Var1', by.x = 'Var2')
donor<-data.frame(table(OL$C2_names, OL$Donor_ID))
donor<-pivot_wider(donor, names_from = Var1, values_from = Freq)
x<-data.frame(table(data$Donor_ID))
donor<-merge(donor, x, by.y = 'Var1', by.x = 'Var2')

proportion_donor <- donor[, -ncol(donor)]
for (col in colnames(proportion_donor)[-1]) {
proportion_donor[[col]] <- proportion_donor[[col]] / donor$Freq
}
for (col in colnames(proportion_donor)[-1]) {
proportion_donor[[col]] <- proportion_donor[[col]] / donor$Freq
}
test<-pivot_longer(proportion_donor, c('Oligo_Mature_ACSBG1', 'Oligo_Mature_OPALIN', 'Oligo_Precursor_SEZ6L', 'Oligo_Precursor_GPR17'))
a<-ggplot(test, aes(x = name, y = value, fill = Var2)) +
geom_bar(stat = "identity") +
labs(x = "CellType", y = "Proportion", fill = "Donor") +
theme_minimal() +
theme(axis.text.x = element_text(angle = 45, hjust = 1))


proportion_dataset <- dataset[, -ncol(dataset)]
for (col in colnames(proportion_dataset)[-1]) {
proportion_dataset[[col]] <- proportion_dataset[[col]] / dataset$Freq
}
test<-pivot_longer(proportion_dataset, c('Oligo_Mature_ACSBG1', 'Oligo_Mature_OPALIN', 'Oligo_Precursor_SEZ6L', 'Oligo_Precursor_GPR17'))
b<-ggplot(test, aes(x = name, y = value, fill = Var2)) +
geom_bar(stat = "identity") +
labs(x = "CellType", y = "Proportion", fill = "Dataset") +
theme_minimal() +
theme(axis.text.x = element_text(angle = 45, hjust = 1))

percent_dataset<-dataset[, -ncol(dataset)]
col_sums<-colSums(percent_dataset[, -1])
for (i in 2:ncol(percent_dataset)) {
  percent_dataset[, i] <- (percent_dataset[, i] / col_sums[i - 1]) * 100
  }
test<-pivot_longer(percent_dataset, c('Oligo_Mature_ACSBG1', 'Oligo_Mature_OPALIN', 'Oligo_Precursor_SEZ6L', 'Oligo_Precursor_GPR17'))
d<-ggplot(test, aes(x = name, y = value, fill = Var2)) +
geom_bar(stat = "identity") +
labs(x = "CellType", y = "Percent", fill = "Dataset") +
theme_minimal() +
theme(axis.text.x = element_text(angle = 45, hjust = 1))


percent_donor<-donor[, -ncol(donor)]
col_sums<-colSums(percent_donor[, -1])
for (i in 2:ncol(percent_donor)) {
  percent_donor[, i] <- (percent_donor[, i] / col_sums[i - 1]) * 100
  }
test<-pivot_longer(percent_donor, c('Oligo_Mature_ACSBG1', 'Oligo_Mature_OPALIN', 'Oligo_Precursor_SEZ6L', 'Oligo_Precursor_GPR17'))
c<-ggplot(test, aes(x = name, y = value, fill = Var2)) +
geom_bar(stat = "identity") +
labs(x = "CellType", y = "Percent", fill = "donor") +
theme_minimal() +
theme(axis.text.x = element_text(angle = 45, hjust = 1))

a + b + c + d
```


## Fig 3-17
```{r, eval = FALSE}
astros<-subset(data, subset = C1 == c('C1-9', 'C1-10'))
DimPlot(astros, group.by = 'C3', reduction = 'umap_scvi_hypo', raster = F) +
  xlim(-6,3) +
  ylim(-6,3) +
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank())

DimPlot(astros, group.by = 'Donor_ID', reduction = 'umap_scvi_hypo', raster = F) +
  xlim(-6,3) +
  ylim(-6,3) +
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank())

Stacked_VlnPlot(astros, c('GFAP', 'AQP4', 'SLC1A2', 'SLC1A3', 'EPHA6', 'TSHZ2', 'ALDH1A1', 'GABBR2', 'CCDC85A', 'LRRTM4', 'LUCAT1', 'ITPRID1', 'LGR6', 'CPAMD8', 'HILPDA', 'FSTL5', 'SERPINA3', 'MT1H', 'ATRNL1', 'NRXN3', 'S100A1', 'CHI3L1'), group.by = 'C4')

```

## Fig 2-18
```{r, eval = FALSE}
dataset<-data.frame(table(astros$C2, astros$Dataset))
dataset<-pivot_wider(dataset, names_from = Var1, values_from = Freq)
x<-data.frame(table(data$Dataset))
dataset<-merge(dataset, x, by.y = 'Var1', by.x = 'Var2')
donor<-data.frame(table(astros$C2, astros$Donor_ID))
donor<-pivot_wider(donor, names_from = Var1, values_from = Freq)
x<-data.frame(table(data$Donor_ID))
donor<-merge(donor, x, by.y = 'Var1', by.x = 'Var2')

proportion_donor <- donor[, -ncol(donor)]
for (col in colnames(proportion_donor)[-1]) {
proportion_donor[[col]] <- proportion_donor[[col]] / donor$Freq
}
for (col in colnames(proportion_donor)[-1]) {
proportion_donor[[col]] <- proportion_donor[[col]] / donor$Freq
}
test<-pivot_longer(proportion_donor, c('C2-21', 'C2-22', 'C2-23', 'C2-24', 'C2-25'))
a<-ggplot(test, aes(x = name, y = value, fill = Var2)) +
geom_bar(stat = "identity") +
labs(x = "CellType", y = "Proportion", fill = "Donor") +
theme_minimal() +
theme(axis.text.x = element_text(angle = 45, hjust = 1))


proportion_dataset <- dataset[, -ncol(dataset)]
for (col in colnames(proportion_dataset)[-1]) {
proportion_dataset[[col]] <- proportion_dataset[[col]] / dataset$Freq
}
test<-pivot_longer(proportion_dataset, c('C2-21', 'C2-22', 'C2-23', 'C2-24', 'C2-25'))
b<-ggplot(test, aes(x = name, y = value, fill = Var2)) +
geom_bar(stat = "identity") +
labs(x = "CellType", y = "Proportion", fill = "Dataset") +
theme_minimal() +
theme(axis.text.x = element_text(angle = 45, hjust = 1))

percent_dataset<-dataset[, -ncol(dataset)]
col_sums<-colSums(percent_dataset[, -1])
for (i in 2:ncol(percent_dataset)) {
  percent_dataset[, i] <- (percent_dataset[, i] / col_sums[i - 1]) * 100
  }
test<-pivot_longer(percent_dataset, c('C2-21', 'C2-22', 'C2-23', 'C2-24', 'C2-25'))
d<-ggplot(test, aes(x = name, y = value, fill = Var2)) +
geom_bar(stat = "identity") +
labs(x = "CellType", y = "Percent", fill = "Dataset") +
theme_minimal() +
theme(axis.text.x = element_text(angle = 45, hjust = 1))


percent_donor<-donor[, -ncol(donor)]
col_sums<-colSums(percent_donor[, -1])
for (i in 2:ncol(percent_donor)) {
  percent_donor[, i] <- (percent_donor[, i] / col_sums[i - 1]) * 100
  }
test<-pivot_longer(percent_donor, c('C2-21', 'C2-22', 'C2-23', 'C2-24', 'C2-25'))
c<-ggplot(test, aes(x = name, y = value, fill = Var2)) +
geom_bar(stat = "identity") +
labs(x = "CellType", y = "Percent", fill = "donor") +
theme_minimal() +
theme(axis.text.x = element_text(angle = 45, hjust = 1))

a + b + c + d

dataset<-data.frame(table(astros$C3, astros$Dataset))
dataset<-pivot_wider(dataset, names_from = Var1, values_from = Freq)
x<-data.frame(table(data$Dataset))
dataset<-merge(dataset, x, by.y = 'Var1', by.x = 'Var2')
donor<-data.frame(table(astros$C3, astros$Donor_ID))
donor<-pivot_wider(donor, names_from = Var1, values_from = Freq)
x<-data.frame(table(data$Donor_ID))
donor<-merge(donor, x, by.y = 'Var1', by.x = 'Var2')

proportion_donor <- donor[, -ncol(donor)]
for (col in colnames(proportion_donor)[-1]) {
proportion_donor[[col]] <- proportion_donor[[col]] / donor$Freq
}
for (col in colnames(proportion_donor)[-1]) {
proportion_donor[[col]] <- proportion_donor[[col]] / donor$Freq
}
test<-pivot_longer(proportion_donor, c('C3-47', 'C3-48', 'C3-49', 'C3-50', 'C3-51', 'C3-52', 'C3-53'))
a<-ggplot(test, aes(x = name, y = value, fill = Var2)) +
geom_bar(stat = "identity") +
labs(x = "CellType", y = "Proportion", fill = "Donor") +
theme_minimal() +
theme(axis.text.x = element_text(angle = 45, hjust = 1))


proportion_dataset <- dataset[, -ncol(dataset)]
for (col in colnames(proportion_dataset)[-1]) {
proportion_dataset[[col]] <- proportion_dataset[[col]] / dataset$Freq
}
test<-pivot_longer(proportion_dataset, c('C3-47', 'C3-48', 'C3-49', 'C3-50', 'C3-51', 'C3-52', 'C3-53'))
b<-ggplot(test, aes(x = name, y = value, fill = Var2)) +
geom_bar(stat = "identity") +
labs(x = "CellType", y = "Proportion", fill = "Dataset") +
theme_minimal() +
theme(axis.text.x = element_text(angle = 45, hjust = 1))

percent_dataset<-dataset[, -ncol(dataset)]
col_sums<-colSums(percent_dataset[, -1])
for (i in 2:ncol(percent_dataset)) {
  percent_dataset[, i] <- (percent_dataset[, i] / col_sums[i - 1]) * 100
  }
test<-pivot_longer(percent_dataset, c('C3-47', 'C3-48', 'C3-49', 'C3-50', 'C3-51', 'C3-52', 'C3-53'))
d<-ggplot(test, aes(x = name, y = value, fill = Var2)) +
geom_bar(stat = "identity") +
labs(x = "CellType", y = "Percent", fill = "Dataset") +
theme_minimal() +
theme(axis.text.x = element_text(angle = 45, hjust = 1))


percent_donor<-donor[, -ncol(donor)]
col_sums<-colSums(percent_donor[, -1])
for (i in 2:ncol(percent_donor)) {
  percent_donor[, i] <- (percent_donor[, i] / col_sums[i - 1]) * 100
  }
test<-pivot_longer(percent_donor, c('C3-47', 'C3-48', 'C3-49', 'C3-50', 'C3-51', 'C3-52', 'C3-53'))
c<-ggplot(test, aes(x = name, y = value, fill = Var2)) +
geom_bar(stat = "identity") +
labs(x = "CellType", y = "Percent", fill = "donor") +
theme_minimal() +
theme(axis.text.x = element_text(angle = 45, hjust = 1))

a + b + c + d

dataset<-data.frame(table(astros$C4, astros$Dataset))
dataset<-pivot_wider(dataset, names_from = Var1, values_from = Freq)
x<-data.frame(table(data$Dataset))
dataset<-merge(dataset, x, by.y = 'Var1', by.x = 'Var2')
donor<-data.frame(table(astros$C4, astros$Donor_ID))
donor<-pivot_wider(donor, names_from = Var1, values_from = Freq)
x<-data.frame(table(data$Donor_ID))
donor<-merge(donor, x, by.y = 'Var1', by.x = 'Var2')

proportion_donor <- donor[, -ncol(donor)]
for (col in colnames(proportion_donor)[-1]) {
proportion_donor[[col]] <- proportion_donor[[col]] / donor$Freq
}
for (col in colnames(proportion_donor)[-1]) {
proportion_donor[[col]] <- proportion_donor[[col]] / donor$Freq
}
test<-pivot_longer(proportion_donor, c('C4-116', 'C4-117', 'C4-118', 'C4-119', 'C4-120', 'C4-121', 'C4-122', 'C4-123', 'C4-124', 'C4-125', 'C4-126', 'C4-127', 'C4-128'))
a<-ggplot(test, aes(x = name, y = value, fill = Var2)) +
geom_bar(stat = "identity") +
labs(x = "CellType", y = "Proportion", fill = "Donor") +
theme_minimal() +
theme(axis.text.x = element_text(angle = 45, hjust = 1))


proportion_dataset <- dataset[, -ncol(dataset)]
for (col in colnames(proportion_dataset)[-1]) {
proportion_dataset[[col]] <- proportion_dataset[[col]] / dataset$Freq
}
test<-pivot_longer(proportion_dataset, c('C4-116', 'C4-117', 'C4-118', 'C4-119', 'C4-120', 'C4-121', 'C4-122', 'C4-123', 'C4-124', 'C4-125', 'C4-126', 'C4-127', 'C4-128'))
b<-ggplot(test, aes(x = name, y = value, fill = Var2)) +
geom_bar(stat = "identity") +
labs(x = "CellType", y = "Proportion", fill = "Dataset") +
theme_minimal() +
theme(axis.text.x = element_text(angle = 45, hjust = 1))

percent_dataset<-dataset[, -ncol(dataset)]
col_sums<-colSums(percent_dataset[, -1])
for (i in 2:ncol(percent_dataset)) {
  percent_dataset[, i] <- (percent_dataset[, i] / col_sums[i - 1]) * 100
  }
test<-pivot_longer(percent_dataset, c('C4-116', 'C4-117', 'C4-118', 'C4-119', 'C4-120', 'C4-121', 'C4-122', 'C4-123', 'C4-124', 'C4-125', 'C4-126', 'C4-127', 'C4-128'))
d<-ggplot(test, aes(x = name, y = value, fill = Var2)) +
geom_bar(stat = "identity") +
labs(x = "CellType", y = "Percent", fill = "Dataset") +
theme_minimal() +
theme(axis.text.x = element_text(angle = 45, hjust = 1))


percent_donor<-donor[, -ncol(donor)]
col_sums<-colSums(percent_donor[, -1])
for (i in 2:ncol(percent_donor)) {
  percent_donor[, i] <- (percent_donor[, i] / col_sums[i - 1]) * 100
  }
test<-pivot_longer(percent_donor, c('C4-116', 'C4-117', 'C4-118', 'C4-119', 'C4-120', 'C4-121', 'C4-122', 'C4-123', 'C4-124', 'C4-125', 'C4-126', 'C4-127', 'C4-128'))
c<-ggplot(test, aes(x = name, y = value, fill = Var2)) +
geom_bar(stat = "identity") +
labs(x = "CellType", y = "Percent", fill = "donor") +
theme_minimal() +
theme(axis.text.x = element_text(angle = 45, hjust = 1))

a + b + c + d
```


## Fig 3-21
```{r, eval = FALSE}
ventricle<-subset(data, subset = C2_names == 'NN_1_Ventricular')
DimPlot(ventricle, reduction = 'umap_scvi_hypo', group.by = 'C4_names') +
  xlim(-4,2) +
  ylim(-5,2) +
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank())
Stacked_VlnPlot(ventricle, group.by = 'C4_names', c('SOX2', 'VIM', 'FGF10', 'NES', 'RAX', 'CRYM', 'FRZB', 'DIO2'), x_lab_rotate = 45)
```

## Fig 3-22
```{r, eval = FALSE}
#adding the code from the paper for making the UMAP plot and the heatmap of POMC and AGRP clusters 
Idents(data)<-data@meta.data$C6
pomc<-WhichCells(data, idents = c('C6-278', 'C6-279', 'C6-280'))
agrp<-WhichCells(data, idents = c('C6-260'))
data$pomcagrp <- ifelse(colnames(data) %in% pomc, 'POMC',
                        ifelse(colnames(data) %in% agrp, 'AGRP', 'NO'))
cols = c('POMC' = '#325A9BFF', 'AGRP' = '#B10DA1FF', 'NO' = 'lightgrey')
DimPlot(data, group.by = 'pomcagrp', reduction = 'umap_scvi_hypo', raster = F, cols = cols) &
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), legend.text = element_text(size = 25)) &
  labs(title = element_blank())
#Heatmap of POMC and ARGP clusters 
heatmap_genes<-c('POMC', 'AGRP', 'LEPR', 'WIF1', 'CALCR', 'GABRG1', 'SOX3','SFRP2', 'OTP', 'NPY', 'PCSK1', 'PCSK2', 'NAT1', 'CPE', 'PAM')
subset<-subset(data, idents = c('C6-278', 'C6-279', 'C6-280', 'C6-260'))
#try with average expression of genes 
avex<-AverageExpression(subset, features = heatmap_genes)
#set the order of the clusters
avex_reorder<-avex$RNA[,match(c('C6-278', 'C6-279', 'C6-280', 'C6-260'), colnames(avex$RNA))]
clusters<-data.frame(colnames(avex_reorder))
clusters$gene_cluster <- c('POMC', 'POMC', 'POMC', 'AGRP')
rownames(clusters)<- clusters$colnames.avex_reorder.
colnames(clusters) <- c("Cluster", "Cluster Type")

my_colour = list(
  Cluster = c(`C6-278` = 'yellow', `C6-279` = 'orange', `C6-280` = 'skyblue', `C6-260` = 'orange'),
  `Cluster Type` = c(POMC = '#325A9BFF', AGRP = '#B10DA1FF'))

avex<-t(log(avex_reorder+1, 2))
pheatmap(avex, 
         annotation_row = clusters,
         annotation_colors =  my_colour,
         color = viridis(256), 
         border_color = 'lightgrey', 
         cluster_rows = F, 
         cluster_cols = F, 
         breaks = seq(0,3, length.out = 256), 
         fontsize = 10, 
         gaps_row = 3,
         gaps_col = c(2, 8, 10),
         annotation_legend = FALSE,
         cellwidth = 25, 
         cellheight = 25, 
         angle_col = 90, 
         filename = '~/Dropbox/Thesis/Chapter3//Figs/POMC_AGRP_HEATMAP.png')
```

## Fig 3-23
```{r, eval = FALSE}
FeaturePlot(data, 'MC3R', order = T, reduction = 'umap_scvi_hypo', raster = F, cols = c('lightgrey', '#0A1172')) &
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), legend.text = element_text(size = 15), plot.title = element_text(face = "italic", size = 20)) &
  labs(title = 'MC3R')

clusters<-merge(as.data.frame(table(data$C6)), as.data.frame(table(subset(data, MC3R>0, slot = 'counts')$C6)), by = 'Var1', all.x = T)
clusters[is.na(clusters)] <- 0
clusters$percent <- clusters$Freq.y / clusters$Freq.x * 100
quantile(clusters$percent, c(0.9, 0.95, 0.98, 0.99))
mc3r<-subset(data, MC3R>0, slot = 'counts')
mc3r_cluster<-top_gene_clusters(data, mc3r, 0, 5.9, metadata_name = 'C6')
mc3r_cluster<-reformat_table(mc3r_cluster, num_to_name = num_to_name, data = data, gene = 'MC3R')
mc3r_cluster<-add_marker_genes(mc3r_cluster, mc3r_cluster$Cluster, markers)

mc3r_cluster <- mc3r_cluster[order(mc3r_cluster$Percent_MC3R), ]
mc3r_test<-as.character(mc3r_cluster$Cluster)
mc3r_test <- gsub("-", ".", mc3r_test)

mc3r_cluster_subset<-subset(data, idents = c("C6-125", "C6-279", "C6-280", "C6-243", "C6-283", "C6-212", "C6-291", "C6-196", "C6-288", "C6-289"))
VlnPlot(mc3r_cluster_subset, c('MC3R', 'SATB2', 'GAL', 'GHRH', 'ADCYAP1', 'POMC', 'CARTPT', 'KISS1', 'TAC1',  'TAC3', 'AGRP', 'SLC17A6', 'SLC32A1', 'CHAT', 'SLC5A7', 'SLC18A2', 'TH'), group.by = 'C6', stack = T, flip = T)


```

Fig 3-25
``` {r, eval = FALSE}
FeaturePlot(data, 'MC4R', order = T, reduction = 'umap_scvi_hypo', raster = F, cols = c('lightgrey', '#0A1172')) &
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), legend.text = element_text(size = 15), plot.title = element_text(face = "italic", size = 20)) &
  labs(title = 'MC4R')

clusters<-merge(as.data.frame(table(data$C6)), as.data.frame(table(subset(data, MC4R>0, slot = 'counts')$C6)), by = 'Var1', all.x = T)
clusters[is.na(clusters)] <- 0
clusters$percent <- clusters$Freq.y / clusters$Freq.x * 100
quantile(clusters$percent, c(0.9, 0.95, 0.98, 0.99))
mc4r<-subset(data, MC4R>0, slot = 'counts')
mc4r_cluster<-top_gene_clusters(data, mc4r, 0, 20.7, metadata_name = 'C6')
mc4r_cluster<-reformat_table(mc4r_cluster, num_to_name = num_to_name, data = data, gene = 'MC4R')
mc4r_cluster<-add_marker_genes(mc4r_cluster, mc4r_cluster$Cluster, markers)

mc4r_cluster <- mc4r_cluster[order(mc4r_cluster$Percent_MC4R), ]
mc4r_test<-as.character(mc4r_cluster$Cluster)
mc4r_test <- gsub("-", ".", mc4r_test)

mc4r_cluster_subset<-subset(data, idents= c("C6-402", "C6-88",  "C6-276", "C6-295", "C6-181", "C6-407", "C6-112", "C6-224", "C6-253", "C6-43"))

#looking at neuropeptide/homrone expression
nps<-read.delim('../Reference-tables/neuropeptide_hormone_GO.txt')
np_list<-list(nps$Gene.name)
np_list<-np_list[[1]]
x<-unique(np_list)
x<-x[-46]
np_list<-x
np_list <- np_list[!(np_list %in% c('CGB5', 'CENPS-CORT', 'CGB8', 'CGB3', 'CGB1', 'INS-IGF2'))]

#find out which genes are expressed in at least 10% of 1 cluster
np_inc<-calc_percent_cluster(mc4r_cluster_subset, genes = np_list, metadata_name = 'C6')
np_inc<-np_inc[, -grep("^AvEx_", colnames(np_inc))]
np_inc<-np_inc[, !(colnames(np_inc) %in% colnames(np_inc)[colSums(np_inc) ==0])]
np_inc<-np_inc[, !(colnames(np_inc) %in% colnames(np_inc)[apply(np_inc, 2, max)<25])]
#removing genes that are highly expressed in all clusters
np_inc<-np_inc[, !(colnames(np_inc) %in% c('CHGB', 'FBN1', 'COPA', 'VGF'))]
#reorder the genes based on cluster expression & add relevant NT markers
mc4r_nps<-c('KL', 'MME', 'INHA', 'AMH', 'GNRH1', 'IGF1', 'SST', 'CCK', 'FNDC5',  'ENHO',  'OXT', 'AVP', 'GAL', 'PENK', 'ADCYAP1','METRN', 'TRH', 'STC1', 'SLC17A6', 'SLC32A1', 'GAD1', 'GAD2', 'CHAT', 'SLC5A7', 'TPH2')

VlnPlot(mc4r_cluster_subset, features =mc4r_nps, stack = TRUE, flip = T) + NoLegend()

```

## Fig 3-26 3-27
```{r, eval = FALSE}
FeaturePlot(data, 'GIPR', order = T, reduction = 'umap_scvi_hypo', raster = F, cols = c('lightgrey', '#0A1172')) &
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), legend.text = element_text(size = 20)) &
  labs(title = 'GIPR')
DimPlot(data, reduction = 'umap_scvi_hypo', raster = F, group.by = 'C2_names', label = T) & NoLegend()
Idents(data)<-data@meta.data$C2
x<-subset(data, idents = 'C2-9')
VlnPlot(x, c('GIPR', 'LMX1A', 'SLC17A6', 'SLC32A1', 'TH', 'DDC', 'SLC18A2'), group.by = 'C6', stack = T, flip = T) & NoLegend()

```




